\section{Graduate Certificates}
\label{sec:grad_certs}

\subsection{Admission and graduation requirements}
\label{sec:cert_requirements}

Graduate students enrolled at Michigan State University in any
discipline or college may pursue this graduate certificate.
Furthermore, students can apply for the certificate at any time prior
to receiving their primary degree (either Master’s or PhD), and must
apply for the certificate after taking all the necessary courses.

Graduate students pursuing either the Master of Science in CMSE, the
Doctor of Philosophy in CMSE, or a dual PhD in CMSE and a second
subject \textbf{may not  receive} either the Graduate Certificate in
Computational Modeling or the Graduate Certificate in High Performance
Computing.

It is also possible to pursue one of the CMSE graduate certificates
while not being enrolled in an MSU degree program.  This requires an
application via the MSU Graduate admission website, and only requires the
application form and transcripts from any prior undergraduate or
graduate coursework/degrees (in particular, recommendation letters and
personal/research statements are not required).  Please contact the
\href{mailto:cmsegrad@msu.edu}{CMSE Director of Graduate Studies} if
you are interested in pursuing this opportunity.

In order to obtain this graduate certificate the student must have at
least a 3.0 average in the courses that are applied to the
certificate.  Courses where the student has received a grade below 2.5
\textbf{may not apply} toward the requirements of either graduate
certificate.  In addition, students must be in good academic standing.

To apply for this certificate, please consult the
\href{https://cmse.msu.edu/academics/graduate-program/}{CMSE website}
for the application information.  You will need the signature of your
academic advisor as well as the \href{mailto:cmsegrad@msu.edu}{CMSE
Director of Graduate Studies}.

%\newpage

\input{model_cert}

%\newpage

\input{hpc_cert}