\section{Master of Science in CMSE}
\label{sec:ms}

\noindent \textbf{Note about the M.S. program:} At present, admission
directly to the Master of Science degree is only available in
exceptional circumstances.  Students interested in pursuing a graduate
degree in Computational Mathematics, Science and Engineering should
apply directly to the PhD program.  Please contact the
\href{mailto:cmsegrad@msu.edu}{CMSE Director of Graduate Studies} with
any questions.


\subsection{Program overview}

The overall goal of the Master of Science program in CMSE is to give students broad and
deep knowledge of the fundamental techniques used in computational
modeling and data science, as well as significant exposure to at least
one application domain.

\vspace{2mm}
\noindent
Students that have completed this M.S. program will be able to:

\begin{itemize}
\item  Analyze problems in terms of the algorithms and pre-existing
  computational tools required to solve a range of problems in
  computational and data science, and write programs to efficiently
  solve the problem using cutting-edge computational hardware.  

\item  Construct and implement models and simulations of physical,
  biological, and social situations, and use these models/simulations
  to understand experimental or observational data.  

\item  Apply discipline-focused or methodology-focused topics in
  computational and data science to solve problems in the student's
  application domain of choice.

%\item  \red{\textbf{OTHER THINGS HERE?}}

\end{itemize}

\subsection{Program requirements}

Students pursuing a M.S. in CMSE must complete a minimum of 30 credits
of graduate coursework, at least 18 credits of which must be CMSE
courses. There are two possible `tracks' in the program: Plan A, which
includes a thesis, or Plan B, which is entirely coursework-based. The
student's program of study must be approved by the student's guidance
committee within one semester of entering the graduate program. The
student must meet the requirements specified below:

\vspace{3mm}
\noindent
\textbf{Common Requirements for Plan A and Plan B:}

\begin{enumerate}

\item Complete a minimum of three of the following four core CMSE
  courses \textbf{with an average grade of at least 3.3 on the
    corresponding subject exams and no grade lower than 3.0} (at least
  9 credits):

\begin{itemize}
    \item  CMSE 820, Mathematical Foundations of Data Science (3 credits)  
    \item  CMSE 821, Numerical Methods for Differential Equations (3 credits)  
    \item  CMSE/CSE 822, Parallel Programming (3 credits)  
    \item  CMSE 823, Numerical Linear Algebra, I (3 credits)  
\end{itemize}

\item Complete a minimum of 12 credits in complementary coursework
  chosen in consultation with, and approved by, the student's academic
  advisor.  (Note: the academic advisor for M.S. students is typically the CMSE Graduate Director.)
 
\item All M.S. students must complete Responsible Conduct of Research (RCR)
  Training and submit annual reports as specified by the
  \href{https://www.egr.msu.edu/academics/graduate/rcr}{College of
    Engineering}.  See Appendix~\ref{sec:rcr_requirements} for a
  detailed explanation of RCR requirements and a timeline.

\end{enumerate}

\vspace{3mm}
\noindent
\textbf{Additional Requirements for Plan A:} In addition to
requirements 1-3 above, the student must complete a thesis based on
original research on a problem in computational and/or data
science. The student will enroll in a minimum of 4 and a maximum of 8
credits of CMSE 899 (Master's Thesis Research). This thesis research
will culminate in a written thesis to be submitted to, and accepted
by, a guidance committee. 

\vspace{3mm}
\noindent
\textbf{Additional Requirements for Plan B:} Requirements 1-3 will
apply. In lieu of pursuing original research in computational and/or
data science, the student will enroll in additional coursework. This
coursework may include up to 3 credits of CMSE 891 (Independent Study)
if approved by the student's academic advisor. 


\subsubsection{Selection of thesis advisor}

Any faculty member with a non-0\% percentage appointment in the
Department of Computational Mathematics, Science and Engineering may
serve as an advisor for a Master's thesis for a student pursuing Plan
A.  With the permission of the CMSE Graduate Director and CMSE
Graduate Studies Committee, a faculty member without an appointment in
CMSE may serve as a Master's thesis advisor.  

The thesis advisor should be
chosen as soon as possible after the student has decided to pursue the
Plan A degree,  ideally no later than the beginning of the student's second
semester in the graduate program.  At the latest, an advisor should be selected prior to
enrolling in CMSE 899 (Master's Thesis Research).  


\subsubsection{Formation of the guidance committee}
\label{sec:ms_guidance_comm}

The purpose of the M.S. guidance committee is primarily to provide
advice to students about coursework selection and professional
development, and to assist them in the development of their
\href{https://www.egr.msu.edu/grs/}{M.S. Program Plan}, which must be
created during the first semester the student is enrolled in the
program.  In the case of students pursuing a Plan A Master's degree,
the guidance committee is also responsible for advising the student in
their choice of research topic, implementation, analysis, and
presentation of said topic, and accepting the written thesis.

All students pursuing a Master's degree in CMSE will be assigned a
guidance committee upon entrance to the degree program.  For students
pursuing a Plan A degree, their guidance committee is composed of
their graduate thesis advisor.  Students pursuing a Plan B degree will
be assigned a CMSE faculty member as the sole member of their guidance
committee.  This faculty member will be either the CMSE Graduate
Director or a member of the graduate studies committee that has been
chosen because their research expertise complements the student's
educational and career goals.  Students
engaged in either track of the Master's degree are \textbf{strongly encouraged}
to collaborate with their guidance committee to create an
\href{http://caffe.grd.msu.edu/IDP}{Individual Development Plan},
which is a document that provides a planning process that helps
students to identify professional development needs, career
objectives, and to facilitate communication between mentees and their
mentors.  The Graduate School
\href{https://grad.msu.edu/prep}{provides resources} for developing an
\href{http://caffe.grd.msu.edu/IDP}{Individual Development Plan}, and
students are encouraged to consult these resources prior to meeting
with their guidance committee.

\subsubsection{Thesis defense and final oral examination}

Students pursuing a Plan A thesis submit a written thesis to their
guidance committee for approval two weeks before the final date for
thesis deposition during their last semester in the Master's degree
program.  This thesis must  be approved by their thesis advisor prior
to submission to the entire committee.  (Note: this date can be found in the
\href{https://reg.msu.edu/ROInfo/Calendar/Academic.aspx}{MSU Academic
Calendar}, but is typically the Monday after the end of final exam
week.)  Submission should be electronic, and preferably via PDF.
Students who wish to use \LaTeX should use the
\href{http://ctan.org/pkg/msu-thesis}{MSU \LaTeX\ thesis class}, which
will ensure that their thesis conforms to the
\href{https://grad.msu.edu/etd/formatting-guide}{Graduate School
Thesis Formatting Guide}.

The guidance committee will provide writtennte feedback on the thesis within one
week, and the student is expected to make any necessary revisions and
get approval from their guidance committee in a timely manner.  After
their guidance committee approves of the thesis, it must be submitted,
along with an Approval Form, to the Graduate School.  Complete
instructions \href{https://grad.msu.edu/etd}{can be found here}.

No oral examination is required for the Plan A Master's degree.  If a
student requests one, the same procedure as is used for the PhD oral
defense shall be used.  See Section~\ref{sec:phd_dissertation} for
additional details.